The Evolution of Phenotypically Invariant Genetic Networks

### 2017-08-11 kalman_walk_paper.tex
more title ideas
started stuff on hybrid vigor

\section*{Gene regulatory network drift and speciation occurs rapidly under neutrality}
{\centering
Joshua S. Schiffman$^{\dagger}$ \qquad Peter L. Ralph$^{\dagger \ddagger}$ \\
$^{\dagger}$University of Southern California, Los Angeles, California \qquad 
$^{\ddagger}$University of Oregon, Eugene, Oregon \\
\texttt{jsschiff@usc.edu} \qquad 
\texttt{plr@uoregon.edu}
\\
}

% Other title ideas:
% More than one way to grow a cat: regulatory network redundancy and speciation
% 
% Neutral network drift leads to incompatibilities on the same time scale as genetic drift
% 
% More than one way to grow a cat: can network drift lead to speciation?
% 
% How fast does network drift create incompatibilities?
% 
% Beyond the snowball: a quantitative model of incompatibility accumulation
% 
% Systems drift and speciation: more than one way to grow a cat
% 
% An explicit model of neutral regulatory network evolution, with applications
% to the rate of accumulation of hybrid incompatibility
%

@@

\section{Linear Evolution-Variant Systems}
  Let $V(\tau) = \prod_{\ell = 0}^{\tau} (I + Q_\ell)$ (where $Q_\ell$ is a random matrix with norm $< \sigma$), and let $A$ be minimal. Then the system architecture after evolutionary time $\tau$ is given by,
    \begin{align}
      \mathcal{S}(\tau) := \left\{ \begin{array}{ll} V(\tau)\dot{\kappa}(t) &= V(\tau)A V(\tau)^{-1} \kappa(t) + V(\tau)B u(t) \\ \phi(t) &= C V(\tau)^{-1} \kappa(t) \end{array} \right .
  \end{align}
\section{Genetic drift with a multivariate trait}
\label{ss:quant_gen}

@@

>>>>

\section*{Rapid speciation despite conservation of phenotype}
{\centering
Joshua S. Schiffman$^{\dagger}$ \qquad Peter L. Ralph$^{\dagger \ddagger}$ \\
$^{\dagger}$University of Southern California, Los Angeles, California \qquad 
$^{\ddagger}$University of Oregon, Eugene, Oregon \\
\texttt{jsschiff@usc.edu} \qquad 
\texttt{plr@uoregon.edu}
\\
}

% Other title ideas:
%
% More than one way to grow a cat: regulatory network drift and speciation
% 
% Rapid speciation despite conservation of phenotype
%
% More than one way to grow a cat: regulatory network redundancy and speciation
% 
% Neutral network drift leads to incompatibilities on the same time scale as genetic drift
% 
% More than one way to grow a cat: can network drift lead to speciation?
% 
% How fast does network drift create incompatibilities?
% 
% Beyond the snowball: a quantitative model of incompatibility accumulation
% 
% Systems drift and speciation: more than one way to grow a cat
% 
% An explicit model of neutral regulatory network evolution, with applications
% to the rate of accumulation of hybrid incompatibility
% 
% Evolutionary conservation of phenotype does not entail conservation of the underlying
% molecular mechanism leading to rapid speciation
%
% The evolution of phenotype-invariant gene networks rapidly leads to hybrid incompatibiliy
%
% Evolutionary network rewiring can rapidly lead to hybrid incompatibility despite
% phenotypic conservation
%
% Hybrid incompatibilities can evolve rapidly due to developmental systems drift
%
% Evolutionary systems theory and speciation

@@

\section{Genetic drift with a multivariate trait}
\label{ss:quant_gen}

@@

\section{Hybrid Vigor}

If a fitness optimum is at the origin $(0,0) \in \R^{2}$, and two populations have drifted away from the optimum by an distance of $r$, then the expected value of their $F_1$ hybrids is,

\begin{align*}
  \mathbb{E}\left[ \left\lVert \frac{1}{2} ((x_1,y_1) + (x_2,y_2))  \right\rVert \right] = \\
  \frac{\sqrt{2}}{8}\left(\sqrt{3\, r^2 - {\sin^{-1}\!\left(r\right)}^2 + r^4 - 2\, r\, \sin^{-1}\!\left(r\right)\, \sqrt{1 - r^2}} + \sqrt{{\sin^{-1}\!\left(r\right)}^2 + 5\, r^2 - r^4 + 2\, r\, \sin^{-1}\!\left(r\right)\, \sqrt{1 - r^2}}\right)
\end{align*}

If $r < \sim 5.5$ then we will observe hybrid vigor, however if $r$ is sufficiently large, the hybrids will have lower fitness, on average.

Some $F_2$s will have perfect fitness, and some will be worse off than both parents. \jss{I still need to do the calculations.} For instance if $r=1$ so that parental populations reside on the unit circle, two randomly drifted populations will create $F_1$s a distance of, on average, $0.6684$. Some $F_2$s from this population wil be back at the origin (a quarter of the time), others as far away as $\sqrt{2}$ (a quarter of the time), and some will stay a distance of $1$ away (half the time).
